---
title: "Negbin GP Regression"
output: html_notebook
---

```{r}
library(cmdstanr)
library(posterior)
library(dplyr)
library(colormap)
library(ggplot2)
library(bayesplot)
```

```{r}
check_cmdstan_toolchain()
```

```{r}
set_cmdstan_path("C:\\stan\\cmdstan-2.26.1")
```

```{r}
cmdstan_path()
```


```{r}
cmdstan_version()
```

Set up some colours

```{r}
c_light <- c("#DCBCBC")
c_light_highlight <- c("#C79999")
c_mid <- c("#B97C7C")
c_mid_highlight <- c("#A25050")
c_dark <- c("#8F2727")
c_dark_highlight <- c("#7C0000")

nom_colors <- c("#DCBCBC", "#C79999", "#B97C7C", "#A25050", "#8F2727", "#7C0000")

c_light_teal <- c("#6B8E8E")
c_mid_teal <- c("#487575")
c_dark_teal <- c("#1D4F4F")
```

Read in the data we generated using SimulateNegBinData.R.

```{r}
data = read.csv('../Outputs/SimulatedNegBinData.csv')
```

Plot the data to see what we're working with

```{r}
ggplot(data = data, aes(x = time_of_observation, y = y_obs)) + 
  geom_jitter(alpha = 0.3) +
  ggtitle('Observed negative binomial process') + 
  xlab('Time') + 
  ylab('Count')
```

# Generate some gaussian processes

Load in a stan model that can simulate a Gaussian process

```{r}
simulate_gaussian_process = cmdstan_model('../StanPrograms/SimGP.stan')
```

```{r}
simulate_gaussian_process$print()
```

```{r}
data_in = list(
  T = 100,
  x = 1:100,
  rho = 2,
  alpha = 1
)

simulated_gp = simulate_gaussian_process$sample(
  data = data_in,
  fixed_param = TRUE,
  chains = 1,
  seed = 1
)
```

Extract the simulated data

```{r}
draws_df = as_draws_df(simulated_gp$draws())

draws_f = draws_df %>%
  select(starts_with('f['))

draws_squared_f = draws_df %>%
  select(starts_with('squared_f['))
```

Plot some observations of the Gaussian process

```{r}
qplot(1:100, as.numeric(draws_f[1,]), geom = 'line') +
  ggtitle("Gaussian process draw") +
  ylab('y') +
  xlab('Time')
```

# fit a negative binomial with a gaussian process prior on the mean and overdispersion

```{r}
fit_negbin_gaussian_process_priors = cmdstan_model('../StanPrograms/fit_negbin_gaussian_process_priors.stan')
```

```{r}
fit_negbin_gaussian_process_priors$print()
```

setup for passing data to stan

```{r}
x_obs = data$time_of_observation
X = model.matrix( ~ as.factor(x_obs) - 1)

data_in = list(
  T = 100,
  N = dim(data)[1],
  X = X,
  y_obs = data$y_obs,
  x = 1:100
)
```

Start running the model fit. This will probably print out a load of messages, this is expected. It's due to the nature of proposing highly constrained variables, such as covariance matrices (which we do a lot of). It should settle down once it starts mixing well, but is really annoying.

```{r}
gp_fit = fit_negbin_gaussian_process_priors$sample(
  data = data_in,
  seed = 1,
  parallel_chains = 4
)
```

```{r}

```


```{r}
gp_fit_draws <- gp_fit$draws()
summarise_draws(subset(gp_fit_draws,
                       variable=c('mu_gp_mean',
                                  'phi_gp_mean',
                                  'rho_mu',
                                  'alpha_mu',
                                  'rho_phi',
                                  'alpha_phi'),
                       regex=TRUE))
```




```{r}
launch_shinystan(gp_fit)

samples = rstan::extract(simu)
```

```{r}
y_obs = samples$y[5,]
qplot(1:100, y_obs, alpha = I(0.2))
```



```{r}
probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
cred <- sapply(1:100,
               function(n) quantile((samples$y_predict[,n]), probs=probs))

x_vals = data_in$x

plot(1, type="n", main='credible intervals',
     xlab="x", ylab="f", xlim=c(min(x_vals), max(x_vals)), ylim = c(0, 70))


polygon(c(x_vals, rev(x_vals)), c(cred[1,], rev(cred[9,])),
        col = c_light, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[2,], rev(cred[8,])),
        col = c_light_highlight, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[3,], rev(cred[7,])),
        col = c_mid, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[4,], rev(cred[6,])),
        col = c_mid_highlight, border = NA)
lines(x_vals, cred[5,], col=c_dark, lwd=2)
```


```{r}

probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
cred <- sapply(1:100,
               function(n) quantile(exp(samples$mu_predict[,n]), probs=probs))

x_vals = data_in$x

plot(1, type="n", main='credible intervals',
     xlab="x", ylab="f", xlim=c(min(x_vals), max(x_vals)), ylim = c(-20, 20))


polygon(c(x_vals, rev(x_vals)), c(cred[1,], rev(cred[9,])),
        col = c_light, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[2,], rev(cred[8,])),
        col = c_light_highlight, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[3,], rev(cred[7,])),
        col = c_mid, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[4,], rev(cred[6,])),
        col = c_mid_highlight, border = NA)
lines(x_vals, cred[5,], col=c_dark, lwd=2)
```

```{r}
gp_fit$cmdstan_diagnose()
```


```{r}

probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
cred <- sapply(1:100,
               function(n) quantile(exp(samples$phi_predict[,n]), probs=probs))

x_vals = data_in$x

plot(1, type="n", main='credible intervals',
     xlab="x", ylab="f", xlim=c(min(x_vals), max(x_vals)), ylim = c(1, 7))


polygon(c(x_vals, rev(x_vals)), c(cred[1,], rev(cred[9,])),
        col = c_light, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[2,], rev(cred[8,])),
        col = c_light_highlight, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[3,], rev(cred[7,])),
        col = c_mid, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[4,], rev(cred[6,])),
        col = c_mid_highlight, border = NA)
lines(x_vals, cred[5,], col=c_dark, lwd=2)
```
```

