---
title: "Generates negative binomial data"
output: html_notebook
---

```{r}
library(ggplot2)
library(colormap)
library(cmdstanr)
library(posterior)
library(dplyr)
```

```{r}
# we recommend running this is a fresh R session or restarting your current session
#("cmdstanr", repos = c("https://mc-stan.org/r-packages/", getOption("repos")))
```


```{r}
# Installing to my cloud backup drive didn't work
#install_cmdstan(dir = 'C:\\stan')
#rebuild_cmdstan()

# check everything is fine with the compilation toolchain
check_cmdstan_toolchain()

# if necessary, install cmdstan
```

```{r}
cmdstan_path()
```

```{r}
cmdstan_version()
```


Set up some default colours for use later.

```{r}

c_light <- c("#DCBCBC")
c_light_highlight <- c("#C79999")
c_mid <- c("#B97C7C")
c_mid_highlight <- c("#A25050")
c_dark <- c("#8F2727")
c_dark_highlight <- c("#7C0000")

nom_colors <- c("#DCBCBC", "#C79999", "#B97C7C", "#A25050", "#8F2727", "#7C0000")

c_light_teal <- c("#6B8E8E")
c_mid_teal <- c("#487575")
c_dark_teal <- c("#1D4F4F")
```

# Generate the mean and overdispersion processes 

We generate some data, where the mean and overdispersion vary over time.

Both the mean and overdispersion are modelled using the exponent of a random walk process (since we want strictly positive values. We then plug these into a Stan model to generate some negative binomial count data. Since we use a Stan model to estimate the underlying process later, it makes sense to generate the negative binomial data using a Stan model, so that the parameterisation of the negative binomial doesn't change.

```{r}
set.seed(1)
T = 100                 # length of time we observe the process over
x = seq(1, T)           # 1,2,3...,T
mu = exp(2 + cumsum(rnorm(T, 0, 0.05))) # exponentiated random walk
phi = exp(cumsum(rnorm(T, 0, 0.15)))    # exponentiated random walk
```

## Mean process

```{r}
qplot(x, mu) + ggtitle("Mean over time") + xlab("Time")
```

# Overdispersion process

```{r}
qplot(x, phi) + ggtitle("phi over time") + xlab("Time")
```

# observations per time point

We allow multiple observations of the negative binomial process at each time point. Each day, the number of observations is drawn from a poisson(10) distribution

```{r}
obs_each_day = rpois(T, 10)
```

```{r}
qplot(1:T, obs_each_day) + 
  ggtitle('Number of observations of the process made at each time point') +
  xlab('Time')
```

We create a design matrix

```{r}
time_of_observation = rep(1:T, obs_each_day)
X = model.matrix( ~ as.factor(time_of_observation) - 1) # design matrix
N = length(time_of_observation) # total number of observations
```

We write a stan program that can generate the negative binomial data. This command will recompile if needed.

```{r}
generate_negative_binomial_process = cmdstan_model('../StanPrograms/SimNegBinProcess.stan')
```

Print the model code so you can see what it does. Should be straightforward to understand - the design matrix describes which observations occur on which day.

```{r}
generate_negative_binomial_process$print()
```


Set up a list of our inputs

```{r}
data_in = list(
  mu = mu,
  phi = phi,
  T = T,
  N = N,
  X = X
)
```

Run Stan to generate some data

```{r}
simulated_data = generate_negative_binomial_process$sample(
  data = data_in,
  fixed_param = TRUE,
  chains = 1,
  seed = 1
)
```

Extract some outputs. Everything gets returned as a big dataframe 

```{r}
draws_df = as_draws_df(simulated_data$draws())

draws_y_obs = draws_df %>%
  select(starts_with('y_obs'))

draws_y = draws_df %>%
  select(starts_with('y['))
```

Inspect the first observation of this process. We will save this data, and later fit a model to it to try and estimate the mean and overdispersion process.

# Observed data

```{r}
y_obs = as.numeric(draws_y_obs[1,])
qplot(time_of_observation, y_obs, alpha = I(0.2)) +
  ggtitle("Some simulated observations") +
  xlab('Time') +
  ylab('Count')
```

```{r}
simulated_data = data.frame(time_of_observation, y_obs)
```

```{r}
write.csv(simulated_data, file = '../Outputs/SimulatedNegBinData.csv')
```

# Credible intervals

We plot the credible intervals from the neagtive binomial process defined by the mean and overdispersion process we randomly generated above.

```{r}

draws_y_matrix = as.matrix(draws_y)

x_vals = 1:100

probs = c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9)
cred <- sapply(1:100,
               function(n) quantile((draws_y_matrix[,n]), probs=probs))

plot(1, type="n", main='Negative binomial process pdf',
     xlab="x", ylab="f", xlim=c(min(1), max(100)), ylim = c(0, 40))

polygon(c(x_vals, rev(x_vals)), c(cred[1,], rev(cred[9,])),
        col = c_light, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[2,], rev(cred[8,])),
        col = c_light_highlight, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[3,], rev(cred[7,])),
        col = c_mid, border = NA)
polygon(c(x_vals, rev(x_vals)), c(cred[4,], rev(cred[6,])),
        col = c_mid_highlight, border = NA)
lines(x_vals, cred[5,], col=c_dark, lwd=2)
```

